---
title: "serialize_prealloc benchmark"
author: Florian Rupprecht
date: 7/Oct/2022
output: md_document
---

# Experiment: Preallocate buffers in base-R serialization

This is a test implementation of R's serialization memory buffers, that uses `object.size` to preallocate memory buffers.

R-base serialize needs a lot of re-allocations with larger data: 
```{r}
set.seed(pi)

invisible(
  serializeprealloc::serialize_prealloc(runif(1000000), prealloc = FALSE, trace = TRUE)
)
```

## Benchmarks

```{r}
ben <- function(object, num_reps) {
  df <- rbenchmark::benchmark(
    "base-r" = {
      base::serialize(object, NULL)
    },
    "prealloc" = {
      serializeprealloc::serialize_prealloc(object)
    },
    "control" = {
      serializeprealloc::serialize_prealloc(object, prealloc = FALSE)
    },
    replications = num_reps,
    order = "elapsed"
  )
  
  os <- object.size(object)
  df$object_size_kb <- os / 1024
  df$bandwidth_gbps <-
    os * num_reps / 1024 ^ 3 / df$elapsed
  df
}
```

This benchmark tests a variety of object sizes (numeric vectors of size 1 to 10^8)

```{r}
df <- do.call(rbind, lapply(seq(0, 8, by = 0.5), function(i) {
  
  nreps <- if (i < 4) 10000 else if (i < 6) 1000 else if (i < 8) 10 else 1

  ben(runif(10^i), num_reps = nreps)
}))
```

<details>
  <summary>Output dataframe (long)</summary>

```{r}
df |> knitr::kable(row.names = F)
```

</details>

Preallocation is around 3 times faster starting form around 10 MB:

(Control mimics the R-base strategy, but has the same calling conventions and overhead as prealloc)

```{r warning=FALSE}
library(ggplot2)
ggplot(df, aes(df$object_size_kb, df$bandwidth_gbps, color = df$test)) +
  labs(x = "Object size (KB)", y = "Bandwidth (GB/s) (higher is better)") +
  geom_point() +
  geom_line() +
  scale_x_log10()
```

Control shows that prealloc being slower with very small object is caused by calling convention:

```{r warning=FALSE}
ggplot(df, aes(df$object_size_kb, df$relative, color = df$test)) +
  labs(x = "Object size (KB)", y = "Relative performance (lower is better)") +
  geom_point() +
  geom_line() +
  scale_x_log10()
```



Still faster using heavily nested objects:

```{r}
obj <- list()
for (i in seq_len(20)) {
  obj <- list(obj, obj)
}
object.size(obj)
ben(obj, num_reps = 100) |> knitr::kable(row.names = F)
```

Testing Platform

```{r}
R.version
```

Processor	12th Gen Intel(R) Core(TM) i7-12700K, 3610 Mhz, 12 Core(s), 20 Logical Processor(s)

32 GB RAM
